# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r echo=TRUE}
unzip("./RepData_PeerAssessment1/activity.zip")
data <- read.csv("./RepData_PeerAssessment1/activity.csv")
data$date = as.Date(data$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
stepsDay <- aggregate(steps ~ date, data = data, sum, na.rm = TRUE)
head(stepsDay)
str(stepsDay)
hist(stepsDay$steps, xlab="Number of steps taken", main = "Histogram of the total number of steps taken each day")

mean(stepsDay$steps)

median(stepsDay$steps)
```

## What is the average daily activity pattern?
```{r echo=TRUE}
stepsInterval <- aggregate(steps ~ interval, data = data, mean, na.rm = TRUE)
plot(steps ~ interval, data = stepsInterval, type = "l")

stepsInterval[which.max(stepsInterval$steps), ]$interval
```


## Imputing missing values
```{r echo=TRUE}
sum(is.na(data$steps))
```

Replace missing step values with the mean for a given 5-minute interval across the entire observation period

```{r echo=TRUE}
data$steps2 = data$steps
for (i in 1:length(data$steps)) if (is.na(data$steps[i])) {
    data$steps2[i] = mean(data$steps, na.rm = TRUE)
}
```

Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r echo=TRUE}
dataNew = data.frame(steps = data$steps2, date = data$date, interval = data$interval)

stepsDayNew = aggregate(steps ~ date, data = dataNew, sum)

hist(stepsDayNew$steps)

mean(stepsDayNew$steps, na.rm = TRUE)

median(stepsDayNew$steps, na.rm = TRUE)
```
## Are there differences in activity patterns between weekdays and weekends?


```{r echo = TRUE}
day = weekdays(dataNew$date)
dayType = vector()
for (item in day) {
    if (item == "Saturday" || item == "Sunday") {
        dayType = append(dayType, "weekend")
    } else {
        dayType = append(dayType, "weekday")
    }
}
dataNew$dayType = factor(dayType)

avgStepsNew = data.frame(xtabs(steps ~ interval + dayType, aggregate(steps ~ 
    interval + dayType, dataNew, mean)))

library(ggplot2)

qplot(interval, Freq, data = avgStepsNew, facets = dayType ~ .)
```
